Feature relevance via Z-scores
========================================================

```{r}
library("Boruta")
require(data.table)
require(ggplot2)
library("RSQLite")
sqlite <- dbDriver("SQLite")
conn <- dbConnect(sqlite, "localData.db")

# only mass, all T1W features
massQuery <- dbGetQuery(conn, "SELECT *
                        FROM  mass_lesion
                        INNER JOIN lesion ON (mass_lesion.lesion_id = lesion.lesion_id)
                        INNER JOIN f_dynamic ON (lesion.lesion_id = f_dynamic.lesion_id)
                        INNER JOIN f_morphology ON (lesion.lesion_id = f_morphology.lesion_id)
                        INNER JOIN f_texture ON (lesion.lesion_id = f_texture.lesion_id)")

# prune entries and extract feature subsets
# corresponds to 5 entries lesion info, 34 dynamic, 19 morpho, 34 texture fueatures
massfields = names(massQuery[c(0:3,30,31,34:67,70:88,91:114)])
masslesioninfo = massQuery[c(1:3,30,31)]
massfeatures = massQuery[c(1,3,34:67,70:88,91:114)]
massfeatures$BenignNMaligNAnt <- as.factor(massfeatures$BenignNMaligNAnt)
write.csv(massfeatures, file = "allmassfeatures.csv")

# only Nonmass, all T1W features
nonmassQuery <- dbGetQuery(conn, "SELECT *
                          FROM  nonmass_lesion
                          INNER JOIN lesion ON (nonmass_lesion.lesion_id = lesion.lesion_id)
                          INNER JOIN f_dynamic ON (lesion.lesion_id = f_dynamic.lesion_id)
                          INNER JOIN f_morphology ON (lesion.lesion_id = f_morphology.lesion_id)
                          INNER JOIN f_texture ON (lesion.lesion_id = f_texture.lesion_id)")

# prune entries and extract feature subsets
# corresponds to 5 entries lesion info, 34 dynamic, 19 morpho, 34 texture fueatures
nonmassfields = names(nonmassQuery[c(0:3,30,31,34:67,70:88,91:114)])
nonmasslesioninfo = nonmassQuery[c(1:3,30,31)]
nonmassfeatures = nonmassQuery[c(1,3,34:67,70:88,91:114)]
nonmassfeatures$BenignNMaligNAnt <- as.factor(nonmassfeatures$BenignNMaligNAnt)
write.csv(nonmassfeatures, file = "nonmassfeatures.csv")


# all T1W features
lesionsQuery <- dbGetQuery(conn, "SELECT *
                        FROM  lesion
                        INNER JOIN f_dynamic ON (lesion.lesion_id = f_dynamic.lesion_id)
                        INNER JOIN f_morphology ON (lesion.lesion_id = f_morphology.lesion_id)
                        INNER JOIN f_texture ON (lesion.lesion_id = f_texture.lesion_id)")

# prune entries and extract feature subsets
# corresponds to 5 entries lesion info, 34 dynamic, 19 morpho, 34 texture fueatures
lesionsfields = names(lesionsQuery[c(1,23,24,3,5,27:60,63:81,84:107)])
lesioninfo = lesionsQuery[c(1,23,24,3,5)]
features = lesionsQuery[c(1,23,27:60,63:81,84:107)]

M<-subset(features, lesion_label=="massB" | lesion_label=="massM")
ifelse( M$lesion_label == "massB", "mass", "mass") -> M$lesion_label
N<-subset(features, lesion_label=="nonmassB" | lesion_label=="nonmassM")
ifelse( N$lesion_label == "nonmassB", "nonmass", "nonmass") -> N$lesion_label
massnonmassfeatures = data.frame(rbind(M,N))
colnames(massnonmassfeatures)[2] <- "BenignNMaligNAnt"
massnonmassfeatures$BenignNMaligNAnt <- as.factor(massnonmassfeatures$BenignNMaligNAnt)
write.csv(massnonmassfeatures, file = "massnonmassfeatures.csv")

# BenignvsMalignant T1W features
B<-subset(features, lesion_label=="massB" | lesion_label=="nonmassB")
ifelse( B$lesion_label == "massB", "NC", "NC") -> B$lesion_label
M<-subset(features, lesion_label=="massM" | lesion_label=="nonmassM")
ifelse( M$lesion_label == "massM", "C", "C") -> M$lesion_label
BenignvsMalignantfeatures = data.frame(rbind(B,M))
colnames(BenignvsMalignantfeatures)[2] <- "BenignNMaligNAnt"
BenignvsMalignantfeatures$BenignNMaligNAnt <- as.factor(BenignvsMalignantfeatures$BenignNMaligNAnt)
write.csv(BenignvsMalignantfeatures, file = "BenignvsMalignantfeatures.csv")

# multiclass T1W features
multiclassfeatures = features
colnames(multiclassfeatures)[2] <- "BenignNMaligNAnt"
multiclassfeatures$BenignNMaligNAnt <- as.factor(multiclassfeatures$BenignNMaligNAnt)
write.csv(multiclassfeatures, file = "multiclassfeatures.csv")

#Color codes:
#  c("green", "yellow", "red", "blue"),
#  Confirmed, Tentative, Rejected and shadow.
# Blue boxplots correspond to minimal, average and maximum Z score of a shadow attribute. Red
```


```{r}
#Boruta mass:
set.seed(111)
massBoruta <-Boruta(BenignNMaligNAnt ~., data=massfeatures[,2:79], doTrace=2, ntree=500)
massBoruta

plot(massBoruta)

#Boruta non-mass:
set.seed(111)
nonmassBoruta <-Boruta(BenignNMaligNAnt ~., data=nonmassfeatures[,2:79], doTrace=2, ntree=500)
nonmassBoruta

plot(nonmassBoruta)

#Boruta for mass vs nonmass
set.seed(111)
massnonmassBoruta <-Boruta(BenignNMaligNAnt ~., data=massnonmassfeatures[,2:79], doTrace=2, ntree=500)
massnonmassBoruta

plot(massnonmassBoruta)

#Boruta for benign vs malignant
set.seed(111)
BenignvsMalignantBoruta <-Boruta(BenignNMaligNAnt ~., data=BenignvsMalignantfeatures[,2:79], doTrace=2, ntree=500)
BenignvsMalignantBoruta

plot(BenignvsMalignantBoruta)

#Boruta for multiclass
set.seed(111)
multiclassBoruta <-Boruta(BenignNMaligNAnt ~., data=multiclassfeatures[,2:79], doTrace=2, ntree=500)
multiclassBoruta

plot(multiclassBoruta)

```

Now plot Z-scores:
=======================
```{r latticeex}
#########################################
#######################
library("latticeExtra")
rankingsmass <- massBoruta$ImpHistory
rankingsnonmass <- nonmassBoruta$ImpHistory

confirmedmass_features<-massBoruta$finalDecision[massBoruta$finalDecision == "Confirmed"]
# Confirmed mass features
print(confirmedmass_features)

confirmednonmass_features<-nonmassBoruta$finalDecision[nonmassBoruta$finalDecision == "Confirmed"]
# Confirmed nonmass features
print(confirmednonmass_features)


####### proces Masses (add fist confirmed feature)
cfeature = as.data.frame(confirmedmass_features[1])  
massframe=data.frame(zscores=rankingsmass[is.finite(rankingsmass[,rownames(cfeature)]),rownames(cfeature)])
massframe$MorN = "mass"
massframe$feature = rownames(cfeature)
masszscore_selected <- massframe

nonmassframe=data.frame(zscores=rankingsnonmass[is.finite(rankingsnonmass[,rownames(cfeature)]),rownames(cfeature)])
nonmassframe$MorN = "nonmass"
nonmassframe$feature = rownames(cfeature)
masszscore_selected[nrow(masszscore_selected)+1:nrow(nonmassframe),] <-  nonmassframe

masszscore_ttest = numeric(length(confirmedmass_features))
masszscore_ttest[1] = t.test(as.data.frame(massframe)$zscores, as.data.frame(nonmassframe)$zscores, alternative = "greater", paired=FALSE, var.equal = FALSE)$p.value

pvallabels = character(length(confirmedmass_features))
pvallabels[1] <-  rownames(cfeature)

Medz_scoreMass = character(length(confirmedmass_features))
Medz_scoreMass[1] <-  median(massframe$zscores)
Medz_scoreNonmass = character(length(confirmedmass_features))
Medz_scoreNonmass[1] <-  median(nonmassframe$zscores)

# proces remaining confirmed feature Masses
for (i in 2:length(confirmedmass_features)) {
  cfeature = as.data.frame(confirmedmass_features[i])  
  massframe=data.frame(zscores=rankingsmass[,rownames(cfeature)])
  massframe$MorN = "mass"
  massframe$feature = rownames(cfeature)
  masszscore_selected[nrow(masszscore_selected)+1:nrow(massframe),] <-  massframe
  
  nonmassframe=data.frame(zscores=rankingsnonmass[,rownames(cfeature)])
  nonmassframe$MorN = "nonmass"
  nonmassframe$feature = rownames(cfeature)
  masszscore_selected[nrow(masszscore_selected)+1:nrow(nonmassframe),] <-  nonmassframe
  
  # p-values test
  masszscore_ttest[i] = t.test(massframe$zscores[is.finite(massframe$zscores)], nonmassframe$zscores[is.finite(nonmassframe$zscores)], alternative = "greater", paired=FALSE, var.equal = FALSE)$p.value
  pvallabels[i] <-  rownames(cfeature)
  
  Medz_scoreMass[i] <- median(massframe$zscores)
  Medz_scoreNonmass[i] <- median(nonmassframe$zscores)
}

# format p values and order in dataframe
dt <- data.table(masszscore_ttest)
pval <- dt[, list(pvalue = paste0("pval: ", format.pval(masszscore_ttest,  eps=0.0001)))]
  
pvalframe=data.frame(pvalue=pval)
pvalframe$feature = pvallabels
pvalframe$Medz_scoreMass = Medz_scoreMass
pvalframe$Medz_scoreNonmass = Medz_scoreNonmass
print(pvalframe)

```
Create steam plot with MedZ-scores mass vs. non-mass

```{r}

pvalframe$Medz_scoreMass <- as.numeric(pvalframe$Medz_scoreMass)
pvalframe$Medz_scoreNonmass <- as.numeric(pvalframe$Medz_scoreNonmass)
#pvalframe <- pvalframe[order(pvalframe$Medz_scoreMass, decreasing = TRUE),] # sort decre by Medz_scoreMass
pvalframe$feature <- factor(pvalframe$feature) # it must be a factor
pvalframe$Medz_scoreNonmass[pvalframe$Medz_scoreNonmass==-Inf] = 0.0

pvalframe$color <- "skyblue3"
pvalframe$color[18:19] <- "mediumpurple3"
pvalframe$color[20:27] <- "darkseagreen4"

dotchart(pvalframe$Medz_scoreMass, labels=pvalframe$feature, cex=0.8,
    pch=22, lwd=2, xlim=range(-1,8), 
    gcolor="black", lcolor=pvalframe$color, bg=pvalframe$color,
    main="Median Z-scores for mass and non-mass lesions\ngrouped feature type",
    xlab="median Z-score relevance")
par(new=TRUE)
dotchart(pvalframe$Medz_scoreNonmass, labels=pvalframe$feature, cex=0.8,
          pch=24, lwd=2, xlim=range(-1,8), 
         gcolor="black", lcolor=pvalframe$color, bg=pvalframe$color)
legend("topright", legend=c("mass", "nonmass"), pch=c(22,24), lwd=2, col=c("black", "black"))

```



# generate boxplot comparison of relevant mass features vs. the same non-mass feature

```{r fig.width=18, fig.height=6, echo=FALSE}
######################### NOTE PVALUES PRINTING IS NOT IN ORDER NEED TO MANUAL ADJUST
ggplot() + 
  geom_boxplot(data = masszscore_selected, mapping=aes(x=MorN, y=zscores, fill = factor(MorN))) + 
  geom_text(data = pvalframe, aes(label=pvalue, x="mass", y=9.5), position='dodge') +
  facet_grid(~ feature) + theme_bw(base_size = 16) + 
  labs(title = "Comparison of Z-scores for Mass confirmed features", y="Z-scores")+
  theme(  axis.text.x=element_text(angle=0, face="bold", size=12),
          legend.position = "bottom",
          strip.background = element_rect(fill=NA), 
          panel.grid.major.y = element_line(colour = "black", linetype = "dotted"), 
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_line(colour=NA))

```


```{r}
####### proces Non-Masses
bfeature = as.data.frame(confirmednonmass_features[1])  
nonmassframe=data.frame(zscores=rankingsnonmass[is.finite(rankingsnonmass[,rownames(bfeature)]),rownames(bfeature)])
nonmassframe$MorN = "nonmass"
nonmassframe$feature = rownames(bfeature)
nonmasszscore_selected <- nonmassframe

massframe=data.frame(zscores=rankingsmass[is.finite(rankingsmass[,rownames(bfeature)]),rownames(bfeature)])
massframe$MorN = "mass"
massframe$feature = rownames(bfeature)
nonmasszscore_selected[nrow(nonmasszscore_selected)+1:nrow(massframe),] <-  massframe

nonmasszscore_ttest = numeric(length(confirmednonmass_features))
nonmasszscore_ttest[1] = t.test(as.data.frame(nonmassframe)$zscores, as.data.frame(massframe)$zscores, alternative = "greater", paired=FALSE, var.equal = FALSE)$p.value

Medz_scoreNonmass = character(length(confirmednonmass_features))
Medz_scoreNonmass[1] <-  median(nonmassframe$zscores)
Medz_scoreMass = character(length(confirmednonmass_features))
Medz_scoreMass[1] <-  median(massframe$zscores)

pvallabels = character(length(confirmednonmass_features))
pvallabels[1] <-  rownames(bfeature)

for (i in 2:length(confirmednonmass_features)) {
  bfeature = as.data.frame(confirmednonmass_features[i]) 
  nonmassframe=data.frame(zscores=rankingsnonmass[is.finite(rankingsnonmass[,rownames(bfeature)]),rownames(bfeature)])
  nonmassframe$MorN = "nonmass"
  nonmassframe$feature = rownames(bfeature)
  nonmasszscore_selected[nrow(nonmasszscore_selected)+1:nrow(nonmassframe),] <-  nonmassframe
  
  massframe=data.frame(zscores=rankingsmass[is.finite(rankingsmass[,rownames(bfeature)]),rownames(bfeature)])
  massframe$MorN = "mass"
  massframe$feature = rownames(bfeature)
  nonmasszscore_selected[nrow(nonmasszscore_selected)+1:nrow(massframe),] <-  massframe
  
  # p-values test
  nonmasszscore_ttest[i] = t.test(as.data.frame(nonmassframe)$zscores, as.data.frame(massframe)$zscores, alternative = "greater", paired=FALSE, var.equal = FALSE)$p.value
  
  pvallabels[i] <-  rownames(bfeature)  
  Medz_scoreMass[i] <- median(massframe$zscores)
  Medz_scoreNonmass[i] <- median(nonmassframe$zscores)
}

# format p values and order in dataframe
dt <- data.table(nonmasszscore_ttest)
pval <- dt[, list(pvalue = paste0("pval: ", format.pval(nonmasszscore_ttest,  eps=0.0001)))]
  
pvalframe=data.frame(pvalue=pval)
pvalframe$feature = pvallabels
pvalframe$Medz_scoreNonmass = Medz_scoreNonmass
pvalframe$Medz_scoreMass = Medz_scoreMass
print(pvalframe)

```
Create steam plot with MedZ-scores mass vs. non-mass

```{r}

pvalframe$Medz_scoreNonmass <- as.numeric(pvalframe$Medz_scoreNonmass)
pvalframe$Medz_scoreMass <- as.numeric(pvalframe$Medz_scoreMass)
#pvalframe <- pvalframe[order(pvalframe$Medz_scoreMass, decreasing = TRUE),] # sort decre by Medz_scoreMass
pvalframe$feature <- factor(pvalframe$feature) # it must be a factor
pvalframe$Medz_scoreNonmass[pvalframe$Medz_scoreNonmass==-Inf] = 0.0

pvalframe$color <- "skyblue3"
pvalframe$color[14] <- "mediumpurple3"
pvalframe$color[15:17] <- "darkseagreen4"

dotchart(pvalframe$Medz_scoreNonmass, labels=pvalframe$feature, cex=0.8,
          pch=24, lwd=2, xlim=range(-1,8), 
         gcolor="black", lcolor=pvalframe$color, bg=pvalframe$color)
par(new=TRUE)
dotchart(pvalframe$Medz_scoreMass, labels=pvalframe$feature, cex=0.8,
    pch=22, lwd=2, xlim=range(-1,8), 
    gcolor="black", lcolor=pvalframe$color, bg=pvalframe$color,
    main="Median Z-scores for mass and non-mass lesions\ngrouped by feature type",
    xlab="median Z-score relevance")
legend("topright", legend=c("mass", "nonmass"), pch=c(22,24), lwd=2, col=c("black", "black"))


```

```


```{r fig.width=18, fig.height=6, echo=FALSE}
# generate boxplot comparison of relevant nonmass features vs. the same mass feature
ggplot() + 
  geom_boxplot(data = nonmasszscore_selected, mapping=aes(x=MorN, y=zscores, fill = factor(MorN))) + 
  geom_text(data = pvalframe, aes(label=pvalue, x="mass", y=9.5)) +
  facet_grid(~ feature) + theme_bw(base_size = 16) + 
  labs(title = "Comparison of Z-scores for Non-mass confirmed features", y="Z-scores")+
  theme(  axis.text.x=element_text(angle=0, face="bold", size=12),
          legend.position = "bottom",
          strip.background = element_rect(fill=NA), 
          panel.grid.major.y = element_line(colour = "black", linetype = "dotted"), 
          panel.grid.major.x = element_blank(), 
          panel.grid.minor = element_line(colour=NA))


```


