Feature relevance stage 1
====================
```{r}
setwd("Z:/Cristina/MassNonmass/Section1 - ExperimentsUpToDate/experimentsRadiologypaper-revision/boruta")
read_inputdataStage1 <- function(subdata, ids) {
  library("RSQLite")
  sqlite <- dbDriver("SQLite")
  conn <- dbConnect(sqlite, "stage1localData.db")
  
  # 2) all T1W features
  lesionsQuery <- dbGetQuery(conn, "SELECT *
           FROM  stage1features
           INNER JOIN lesion ON (stage1features.lesion_id = lesion.lesion_id)
           INNER JOIN f_dynamic ON (stage1features.lesion_id = f_dynamic.lesion_id)
           INNER JOIN f_morphology ON (stage1features.lesion_id = f_morphology.lesion_id)
           INNER JOIN f_texture ON (stage1features.lesion_id = f_texture.lesion_id)")
  
  # prune entries and extract feature subsets
  # corresponds to 5 entries lesion info, 34 dynamic, 19 morpho, 34 texture fueatures
  lesionfields =  names(lesionsQuery)
  lesioninfo = lesionsQuery[c(1,2,150,151)]
  stage1features = lesionsQuery[c(3:103,124:127)]
  dynfeatures = lesionsQuery[c(154:187)]
  morphofeatures = lesionsQuery[c(190:208)]
  texfeatures = lesionsQuery[c(211:234)]
  
  # combine all features
  allfeatures = cbind(lesioninfo[c(2,3)], stage1features, dynfeatures, morphofeatures, texfeatures)   
  
  if(subdata=="stage2"){
    # organized the data by subdata
    allfeatures = allfeatures[ids,]
    M<-subset(allfeatures, lesion_label=="massB" | lesion_label=="massM")
    ifelse( M$lesion_label == "massB", "NC", "C") -> M$lesion_label
    N<-subset(allfeatures, lesion_label=="nonmassB" | lesion_label=="nonmassM")
    ifelse( N$lesion_label == "nonmassB", "NC", "C") -> N$lesion_label
    allfeatures = rbind(M,N)
  }
  if(subdata=="stage1"){
    # organized the data by subdata
    allfeatures = allfeatures[ids,]
    M<-subset(allfeatures, lesion_label=="massB" | lesion_label=="massM")
    ifelse( M$lesion_label == "massB", "mass", "mass") -> M$lesion_label
    N<-subset(allfeatures, lesion_label=="nonmassB" | lesion_label=="nonmassM")
    ifelse( N$lesion_label == "nonmassB", "nonmass", "nonmass") -> N$lesion_label
    allfeatures = data.frame(rbind(M,N))     
  }
  if(subdata=="oneshot"){
    # organized the data by subdata
    allfeatures = allfeatures[ids,]
    M<-subset(allfeatures, lesion_label=="massB" | lesion_label=="massM")
    ifelse( M$lesion_label == "massB", "NC", "C") -> M$lesion_label
    N<-subset(allfeatures, lesion_label=="nonmassB" | lesion_label=="nonmassM")
    ifelse( N$lesion_label == "nonmassB", "NC", "C") -> N$lesion_label
    allfeatures = data.frame(rbind(M,N)) 
  }
  # procees data
  allfeatures$lesion_label <- as.factor(allfeatures$lesion_label)
  allfeatures$peakCr_inside <- as.factor(allfeatures$peakCr_inside)
  allfeatures$peakVr_inside <- as.factor(allfeatures$peakVr_inside)
  allfeatures$peakCr_countor <- as.factor(allfeatures$peakCr_countor)
  allfeatures$peakVr_countor <- as.factor(allfeatures$peakVr_countor)
  allfeatures$k_Max_Margin_Grad <- as.factor(allfeatures$k_Max_Margin_Grad)
  allfeatures$max_RGH_mean_k <- as.factor(allfeatures$max_RGH_mean_k)
  allfeatures$max_RGH_var_k <- as.factor(allfeatures$max_RGH_var_k)
  
  output <- allfeatures
  return(output)
}
```

You can also embed plots, for example:
```{r fig.width=7, fig.height=6}
stage1dat = read_inputdataStage1("stage1", 1:409)

library("Boruta")
require(data.table)
require(ggplot2)
set.seed(111)
stage1datnaomit = na.omit(stage1dat)
stage1Boruta <-Boruta(lesion_label ~., data=stage1datnaomit[,2:184], doTrace=2, ntree=1000)
stage1Boruta
plot(stage1Boruta)

```

Now plot Z-scores:
============
```{r latticeex}
#########################################
#######################
library("latticeExtra")
rankingsstage1 <- stage1Boruta$ImpHistory
confirmedstage1_features<-stage1Boruta$finalDecision[stage1Boruta$finalDecision == "Confirmed"]
# Confirmed stage1 features
print(confirmedstage1_features)

####### proces Masses (add fist confirmed feature)
cfeature = as.data.frame(confirmedstage1_features[1])  
stage1frame=data.frame(zscores=rankingsstage1[is.finite(rankingsstage1[,rownames(cfeature)]),rownames(cfeature)])
stage1frame$MorN = "stage1"
stage1frame$feature = rownames(cfeature)
stage1zscore_selected <- stage1frame

Medz_scorestage1 = character(length(confirmedstage1_features))
Medz_scorestage1[1] <-  median(stage1frame$zscores)
featlabels = character(length(confirmedstage1_features))
featlabels[1] <-  rownames(cfeature)

# proces remaining confirmed feature stage1es
for (i in 2:length(confirmedstage1_features)) {
  cfeature = as.data.frame(confirmedstage1_features[i])  
  stage1frame=data.frame(zscores=rankingsstage1[,rownames(cfeature)])
  stage1frame$MorN = "stage1"
  stage1frame$feature = rownames(cfeature)
  stage1zscore_selected[nrow(stage1zscore_selected)+1:nrow(stage1frame),] <-  stage1frame
  
  Medz_scorestage1[i] <- median(stage1frame$zscores)
  featlabels[i] <-  rownames(cfeature)
}

# format 
featframe=data.frame(feature=featlabels)
featframe$Medz_scorestage1 = Medz_scorestage1
print(featframe)
featframe$color <- "mediumpurple3"

#plot
dotchart(as.numeric(featframe$Medz_scorestage1), labels=featframe$feature, cex=0.8,
    pch=22, lwd=2, xlim=range(-1,8), 
    gcolor="black", lcolor=featframe$color, bg=featframe$color,
    main="Median Z-scores for stage1 feature type",
    xlab="median Z-score relevance")

